---
title: "Gray-headed Chickadee Simulation Analysis"
author: "David Iles"
date: "Generated on `r Sys.Date()`"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "png",
  dev.args = list(type = "cairo-png"),
  fig.width = 6,
  fig.height = 4,
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  error = FALSE,
  message = FALSE
)
```

# Motivation

The goal of this study is to evaluate the sample size requirements to estimate population sizes of Gray-headed Chickadee in Alaska.

# Simulation description

```{r, include=FALSE, echo = FALSE}
rm(list=ls())
load("power_wksp.RData")
```

The simulation assumes that bird territories are distributed randomly across available habitat and that bird territory boundaries do not overlap.  These simulations assume the area of potentially suitable GHCH habitat in AK is `r plot_area` km^2.




```{r cars}

```

# Simulation results

You can also embed plots, for example:

```{r pressure, echo=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
