---
title: "Gray-headed Chickadee Simulation Analysis"
author: "David Iles"
date: "Generated on `r Sys.Date()`"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "png",
  dev.args = list(type = "cairo-png"),
  fig.width = 12,
  fig.height = 8,
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  error = FALSE,
  message = FALSE
)
```

# Motivation

The goal of this study is to evaluate the sample size requirements to estimate population sizes of Gray-headed Chickadee in Alaska.

# Simulation description

```{r, include=FALSE, echo = FALSE}
library(sp)
library(sf)
library(tidyverse)
library(jagsUI)

rm(list=ls())
options(scipen=10000)
load("power_wksp.RData")
```

These simulations place home ranges on a hypothetical landscape, and conduct simple random sampling across that landscape to estimate the expected number of detections that would occur under various population/sampling scenarios.

The simulated data are then analyzed with a Bayesian model to produce estimates of population size, which can then be compared to the 'true' (i.e., simulated) numbers of birds.  These estimates of population size are then used to conduct precision and power analyses.

### Simulation assumptions

1) bird territories are distributed randomly across available habitat
2) bird territory boundaries do not overlap
3) the area of potentially suitable GHCH habitat in AK is `r plot_area` km<sup>2</sup> (estimated based on remote sensing)
4) surveys are positioned independently of bird occurrence
5) surveys are not clustered

### Example of simulated landscape and survey

We assume in this example that there are 1000 GHCH pairs distributed across the landscape, such that density is `r round(1000/plot_area,3)` birds/km<sup>2</sup>.  Later, we will conduct a full suite of simulations across multiple hypothetical GHCH population sizes.

The plot below illustrates the spatial distribution of GHCH territories for one realization of the simulation.

```{r chunk1, echo = FALSE,fig.width=6, fig.height=6}

n_males = 1000
n_ARU = 300

# --------------------------------------------------
# Simulate bird territories
# --------------------------------------------------

# Bird territories
bird_centroids <- sample_n(birdgrid,n_males)

# Detectable radius of each location
bird_TD = st_buffer(bird_centroids, bird_TrD)

# Plot bird territories
bird_radius_m = round(bird_Tr*1000)

bird_plot <- ggplot() +
  geom_sf(data = bird_TD,fill = "transparent", col = "blue")+
  scale_fill_manual(values = c("transparent","blue"),na.value = "transparent",name="Detection")+
  coord_sf(xlim = c(-dim,dim), ylim = c(-dim,dim))+
  xlab("km")+ylab("km")+
  ggtitle(paste0("Number of males = ",n_males,"\nBird Territory Radius = ",bird_radius_m," m\nBird EDR = ",bird_EDR*1000," m"))

print(bird_plot)

```

Next, we position a number of ARUs on the landscape.

```{r chunk2, echo = FALSE,fig.width=6, fig.height=6}

# --------------------------------------------------
# Simulate ARU placement
# --------------------------------------------------

# SRS sampling
aru_sf <- sample_n(ARUgrid,n_ARU)

aru_detections <- sf::st_intersects(bird_TD,aru_sf) %>% as.data.frame()

bird_TD$det = "no"

if (nrow(aru_detections) > 0){
  bird_TD$det[aru_detections[,1]] <- "yes"
}

# Plot ARU locations
aru_plot <- ggplot() +
  geom_sf(data = bird_TD,fill = "transparent", col = "blue", linetype = 2)+
  geom_sf(data = aru_sf, shape = 19)+
  geom_sf(data = subset(bird_TD, det == "yes"), fill = "red",col="red")+
  coord_sf(xlim = c(-dim,dim), ylim = c(-dim,dim))+
  xlab("km")+ylab("km")+
  ggtitle(paste0("Number of males = ",n_males,"\nBird Territory Radius = ",bird_radius_m," m\nBird EDR = ",bird_EDR*1000," m"))

print(aru_plot)

# --------------------------------------------------
# Calculate number of ARUs with detections
# --------------------------------------------------
n_ARU_det <- length(unique(aru_detections$col.id))

```

In this example we positioned 300 ARUs, and obtained `r n_ARU_det` detections of GHCH.

# Statistical analysis

We analyze the resultant data (`r n_ARU_det` detections across 300 ARUs) using a Bayesian model.  The model assumes that we know the effective detection radius for Gray-headed Chickadee that are surveyed by ARUs.  The model further assumes that since ARUs are positioned and continuously monitoring for the entire breeding season, if the effective detection radius overlaps a GHCH territory it will detect the species at least once over the course of a season.

The analysis is coded in the JAGS language:

```{r analysis1, eval = FALSE, echo = TRUE}

    model {

      # Moderately informative prior on density (assume it is low, but with enough uncertainty to capture higher densities)
      dens_prior_median <- 0.2
      dens_prior_sd <- 2
      
      dens ~ dlnorm(log(dens_prior_median),pow(dens_prior_sd,-2))
      popsize <- IAO_area * dens
      
      total_area_detectable <- popsize * pi * (territory_radius + EDR)^2
      proportion_detectable <- total_area_detectable/IAO_area
      
      # Likelihood
      y ~ dbin(proportion_detectable, n_ARU)
      
      # Output the priors to evaluate if they look reasonable
      dens_prior ~ dlnorm(log(dens_prior_median),pow(dens_prior_sd,-2))
      popsize_prior <- IAO_area * dens_prior
    }
    
```

In this code, we specific a relatively uninformative prior for density, and must supply as data the area of potentially suitable habitat (IAO_area = `r plot_area` km<sup>2</sup>), number of ARUs we positioned in potentially suitable habitat (n_ARU = `r n_ARU`), and the number of ARUs that detected the species (y = `r n_ARU_det`) after being deployed for the entire season.

In brief, the model attempts to estimate the range of population sizes (and population densities) that would be consistent with the observed number of detections, given the amount of survey effort that was conducted.

```{r jags_results, eval = TRUE, echo = TRUE}

    # Analyze using JAGS
  jags.data <- list(y = n_ARU_det,   # Number of ARUs that detected the species
                    n_ARU = n_ARU,
                    IAO_area = plot_area,
                    territory_radius = bird_Tr,
                    EDR = bird_EDR,
                    pi = pi)
  
  out <- jags(data = jags.data,
              model.file = "ghch_model.jags",
              parameters.to.save = c("dens","popsize","popsize_prior","dens_prior"),n.chains = 3,
              n.thin = 1,
              n.iter = 6000,
              n.burnin = 1000,
              verbose = FALSE)
    
```

In this case, the estimate of number of GHCH pairs is: `r round(out$q50$popsize)` (95% CI = `r round(out$q2.5$popsize)` to `r round(out$q97.5$popsize)`); recall that the true population size in this simulation was `r n_males`.  The model estimates there is a `r 100*round(mean(out$sims.list$popsize < 1000),2)`% chance there are fewer than 1000 pairs in the study area.   

These are only results for a single run of the simulation, and we want to understand how the sampling plan and analysis works *on average* - and how variable the results might be; this single run might be unrepresentative of how things tend to shake out.  We therefore run the simulations many times (e.g., 1000 times) and evaluate the spread of results across those many repeated simulations; this is known as Monte Carlo simulation.  

# Simulation parameters

We run simulations for 3 levels of sampling intensity (150, 300, and 600 ARU survey locations), and for 3 levels of population density (n_males = 500, 1000, and 2000) which corresponds to population densities of `r round(500/plot_area,2)`, `r round(1000/plot_area,2)`, and `r round(2000/plot_area,2)` males/km<sup>2</sup>.

Note that this represents a range of low to medium, but biologically plausible population densities that have been observed elsewhere in the species' range (Scandinavia).  

The goal of these simulations is to confirm that we would likely obtain detections of the species and could estimate population size accurately under realistic survey intensity and low population sizes.


# Simulation results

The plot below illustrates the number of detections that could be expected under various scenarios of sampling intensity and population abundance. This figure illustrates that it is highly likely we would obtain at least one (and often more than one) detection under any of the scenarios.

```{r chunk3, echo=FALSE, echo = FALSE,fig.width=12, fig.height=6}

# --------------------------------------------------
# Plot expected number of detections under each scenario
# --------------------------------------------------

n_det_plot <- ggplot(all_results)+
  geom_histogram(aes(x = n_detections), binwidth = 0.5)+
  facet_grid( study_area_label + n_ARU_label ~ n_male_label)+
  geom_text(data = result_summary,aes(x = 10, y = 150, label = paste0("Pr. > 0 dets = ",signif(result_summary$detected_species,2))),hjust=0, size = 3)+
  theme_bw()+
  xlab("Number of detections")+
  ylab("Frequency (number of simulations)")+
  ggtitle("Summary of GHCH detections across repeated simulations")

print(n_det_plot)
```

The plot below illustrates the mean population estimates for each scenario (across 1000 repeated simulations).  Population estimates are nearly unbiased under all scenarios, and increasingly unbiased under higher sampling intensities.  There is a tendency for the model to slightly over-estimate abundance when sampling intensity is low and when population abundance is low.  Precision of the estimates increases as sampling intensity increases.

```{r chunk4, echo=FALSE}

# --------------------------------------------------
# Plot expected counts under each scenario
# --------------------------------------------------

popest_plot <- ggplot(result_summary)+
  geom_hline(aes(yintercept = n_males), linetype = 2)+
  
  # Plot actual estimate (prior x data)
  geom_errorbar(aes(x = n_ARU,ymin = popsize_est_q025, ymax = popsize_est_q975), width = 0)+
  geom_point(aes(x = n_ARU,y = popsize_est_mean))+
  
  facet_grid( study_area_label ~ n_male_label)+
  theme_bw()+
  xlab("Number of ARUs")+
  ylab("Population Estimate")+
  ggtitle(paste0("Population estimates for different combinations of\nn_ARU and n_males"))

print(popest_plot)

```
